    
```{r}
library(stringr)
library(ggplot2)
library(ggridges)
library(hrbrthemes)
library(ggpubr)
library(envalysis)
library(uwot)
library(Rtsne)
library(parallel)
library(factoextra)
library(viridis)
library(dplyr)
library(protr)
```

In this file we are going to do some analysises of our dataset and create some figures

#Database analysis
First lets load our db
```{r}
dataset <- read.csv("/home/jeppe/Dropbox/internship_TNO/Datasets/hemoDatabases/datasetPeptideCompositionFeatures.csv")
sequences <- as.vector(dataset$Sequence)

trainTest <- readRDS("/home/jeppe/Dropbox/internship_TNO/Datasets/hemoDatabases/trainTestSequences.rds")
train <- dataset[dataset$Sequence %in% trainTest[["train"]],]
train$Sequence <- NULL
test <- dataset[dataset$Sequence %in% trainTest[["test"]],]
test$Sequence <- NULL
```

Let's see how many peptides we started with in each of the three datasets 
(HemoPi, Hlppred and HAPPENN) and with how many peptides we ended up in our final
dataset.

First we need our function again to read fastas 
```{r}
# Function to extract fasta. Inputs are path and pattern
# path - folder where the fasta files are located 
# pattern - pattern of the file name(s)
# Output is peptide sequences as character vector
extractFASTA <- function(path, pattern) {
  # Lists the files
  fastas <-
    list.files(path = path,
               pattern = pattern,
               full.names = TRUE)
  # Read all lines of all files and puts them in a vector
  fastas <- unlist(lapply(fastas, readLines))
  # Remove the names of the sequences, which is always on the first line,
  # then removes doubles
  fastas <- unique(fastas[c(FALSE, TRUE)])
  return(fastas)
}
```

And let's load the datasets we have used and analyse them.

```{r}
# HemoPI sequences, split in pos and negs
posHemopi <- extractFASTA(path = "/home/jeppe/Dropbox/internship_TNO/Datasets/HemoPI/", 
                          pattern = ".*3.*pos.*.fa") 
negHemopi <- extractFASTA(path = "/home/jeppe/Dropbox/internship_TNO/Datasets/HemoPI/",
                          pattern = ".*3.*neg.*.fa")

cat("Amount all HemoPI peps, pos and neg\n")
length(c(negHemopi, posHemopi))
length(posHemopi)
length(negHemopi)


# HLPpred sequences, split in pos and negs
posHlppred <- extractFASTA(path = "/home/jeppe/Dropbox/internship_TNO/Datasets/HLPpred", 
                           pattern = ".*2.*positive.txt")
negHlppred <- extractFASTA(path = "/home/jeppe/Dropbox/internship_TNO/Datasets/HLPpred", 
                           pattern = ".*2.*negative.txt")

cat("Amount all HLPpred peps, pos and neg\n")
length(c(posHlppred, negHlppred))
length(posHlppred)
length(negHlppred)


# Happenn
# Read fasta file
HAPPENN <- readLines("/home/jeppe/Dropbox/internship_TNO/Datasets/HAPPENN/HAPPENN_dataset.fasta")
# Store as dataframe 1 column metadata, 1 column sequence and add colnames
HAPPENN <- as.data.frame(matrix(HAPPENN,ncol =2,byrow = T)) 
colnames(HAPPENN) <- c("Hemolytic", "Sequence") 

# Change non-hemolytic metadata to 0 and hemolytic to 1
# and store as vectors by class
HAPPENN$Hemolytic <- gsub(".*.non-hemolytic", "0", x = HAPPENN$Hemolytic) 
HAPPENN$Hemolytic <- gsub(".*.hemolytic", "1", x = HAPPENN$Hemolytic)

posHappenn <- as.character(HAPPENN[which(HAPPENN$Hemolytic == 1), 2]) 
negHappenn <- as.character(HAPPENN[which(HAPPENN$Hemolytic == 0), 2])

cat("Amount all HAPPENN peps, pos and neg\n")
length(c(posHappenn, negHappenn))
summary(str_length(posHappenn))
  length(posHappenn)
length(negHappenn)
```

Here we can find out how many peptides we started with and how many we ended up 
with after combining everything. I used this for the flowchart figure in the results 

```{r}
# Combine HemoPi and HLPpred and remove duplicates
expPos <- c(posHemopi, posHlppred, posHappenn)
expNeg <- c(negHemopi, negHlppred, negHappenn)

cat("Amount all combined peps, pos and neg\n")
length(c(expPos, expNeg))
length(expPos)
length(expNeg)

expPos <- stringi::stri_unique(expPos)
expNeg <- stringi::stri_unique(expNeg)

cat("Amount all unqiue peps per class, pos and neg\n")
length(c(expPos, expNeg))
length(expPos)
length(expNeg)

# Remove sequences that are in both the combined pos and neg datasets
remove1 <- expNeg %in% expPos
remove2 <- expPos %in% expNeg
expNeg <- expNeg[!remove1] 
expPos <- expPos[!remove2]

cat("Amount of dataset peps, hemo and non-hemo after per class removal\n")
length(c(expPos, expNeg))
length(expPos)
length(expNeg)
```

Also, lets check on how many peptides the HemoPred and HLPpred authors and
HAPPENN authors disagreed.

```{r}
# Amount of peptides both HemoPi and Happenn reviewd and classified
length(Reduce(intersect, list(posHemopi, posHappenn))) +
  length(Reduce(intersect, list(negHemopi, negHappenn)))

# Amount of peptides both HLPpred-fuse and Happenn reviewed and classified
length(Reduce(intersect, list(posHlppred, posHappenn))) +
  length(Reduce(intersect, list(negHlppred, negHappenn)))
# Same

# Amount of peptides they disagreed on
length(Reduce(intersect, list(unique(
  c(posHemopi, posHlppred, posHappenn)
), unique(
  c(negHemopi, negHlppred, negHappenn)
))))

# percentage they disagreed on
  length(Reduce(intersect, list(unique(
    c(posHemopi, posHlppred, posHappenn)
  ), unique(
    c(negHemopi, negHlppred, negHappenn)
  )))) / (length(Reduce(intersect, list(
    posHemopi, posHappenn
  ))) +
    length(Reduce(intersect, list(
      negHemopi, negHappenn
    )))) * 100
  
  # Also, do Hlppred and HemoPi disagree on any peps?
length(Reduce(intersect, list(unique(
  c(posHemopi, posHlppred)
), unique(
  c(negHemopi, negHlppred)
))))

# NO
```


More data, number of hemolytic and non hemolytic peps in main dataset

```{r}
cat("Number of total peps\n")
nrow(dataset)
cat("\nNumber of hemo peps and %\n")
length(which(dataset$Hemolytic == 1))
length(which(dataset$Hemolytic == 1))/nrow(dataset) * 100

cat("\nNumber of nonhemo peps and %\n")
length(which(dataset$Hemolytic == 0))
length(which(dataset$Hemolytic == 0))/nrow(dataset) * 100
```

Now for the train and test set

```{r}
cat("Number of train peps\n")
nrow(train)
cat("\nNumber of hemo peps and %\n")
length(which(train$Hemolytic == 1))
length(which(train$Hemolytic == 1))/nrow(train) * 100

cat("\nNumber of nonhemo pepsand %\n")
length(which(train$Hemolytic == 0))
length(which(train$Hemolytic == 0))/nrow(train) * 100

cat("\nNumber of test peps\n")
nrow(test)
cat("\nNumber of hemo peps and %\n")
length(which(test$Hemolytic == 1))
length(which(test$Hemolytic == 1))/nrow(test) * 100

cat("\nNumber of nonhemo pepsand %\n")
length(which(test$Hemolytic == 0))
length(which(test$Hemolytic == 0))/nrow(test) * 100

```

Number of linear/cyclic 

```{r}
cat("Number of linear and cyclic peps and percentages \n")
length(which(dataset$Intrachainbond == "Linear"))
length(which(dataset$Intrachainbond == "Linear"))/nrow(dataset) * 100
length(which(dataset$Intrachainbond == "Cyclic"))
length(which(dataset$Intrachainbond == "Cyclic"))/nrow(dataset) * 100
```

now for test/train again

```{r}
cat("TEST: Number of linear and cyclic peps and percentages \n")
length(which(train$Intrachainbond == "Linear"))
length(which(train$Intrachainbond == "Linear"))/nrow(train) * 100
length(which(train$Intrachainbond == "Cyclic"))
length(which(train$Intrachainbond == "Cyclic"))/nrow(train) * 100

cat("TRAIN: Number of linear and cyclic peps and percentages \n")
length(which(test$Intrachainbond == "Linear"))
length(which(test$Intrachainbond == "Linear"))/nrow(test) * 100
length(which(test$Intrachainbond == "Cyclic"))
length(which(test$Intrachainbond == "Cyclic"))/nrow(test) * 100
```

Same for endterms

```{r}
cat("C_TERM_MODS: Number of free/amiditation peps and percentages \n")
length(which(dataset$C_term_mod == "Free"))
length(which(dataset$C_term_mod == "Free"))/nrow(dataset) * 100
length(which(dataset$C_term_mod == "Amidation"))
length(which(dataset$C_term_mod == "Amidation"))/nrow(dataset) * 100

cat("N_TERM_MODS: Number of free/acetylatio/formylation peps and percentages \n")
length(which(dataset$N_term_mod == "Free"))
length(which(dataset$N_term_mod == "Free"))/nrow(dataset) * 100
length(which(dataset$N_term_mod == "Acetylation"))
length(which(dataset$N_term_mod == "Acetylation"))/nrow(dataset) * 100
length(which(dataset$N_term_mod == "Formylation"))
length(which(dataset$N_term_mod == "Formylation"))/nrow(dataset) * 100
```

now for test/train again

```{r}
cat("TRAIN: C_TERM_MODS: Number of free/amiditation peps and percentages \n")
length(which(train$C_term_mod == "Free"))
length(which(train$C_term_mod == "Free"))/nrow(train) * 100
length(which(train$C_term_mod == "Amidation"))
length(which(train$C_term_mod == "Amidation"))/nrow(train) * 100

cat("TRAIN: N_TERM_MODS: Number of free/acetylatio/formylation peps and percentages \n")
length(which(train$N_term_mod == "Free"))
length(which(train$N_term_mod == "Free"))/nrow(train) * 100
length(which(train$N_term_mod == "Acetylation"))
length(which(train$N_term_mod == "Acetylation"))/nrow(train) * 100
length(which(train$N_term_mod == "Formylation"))
length(which(train$N_term_mod == "Formylation"))/nrow(train) * 100

cat("TEST: C_TERM_MODS: Number of free/amiditation peps and percentages \n")
length(which(test$C_term_mod == "Free"))
length(which(test$C_term_mod == "Free"))/nrow(test) * 100
length(which(test$C_term_mod == "Amidation"))
length(which(test$C_term_mod == "Amidation"))/nrow(test) * 100

cat("TEST: N_TERM_MODS: Number of free/acetylatio/formylation peps and percentages \n")
length(which(test$N_term_mod == "Free"))
length(which(test$N_term_mod == "Free"))/nrow(test) * 100
length(which(test$N_term_mod == "Acetylation"))
length(which(test$N_term_mod == "Acetylation"))/nrow(test) * 100
length(which(test$N_term_mod == "Formylation"))
length(which(test$N_term_mod == "Formylation"))/nrow(test) * 100
```

Now it is nice to visualise our sequence length distributions

```{r}
# Make df with length and class
df <- data.frame("Length" = str_length(sequences), 
                 "Hemolytic" = as.factor(dataset$Hemolytic)) 
# Mean of positive and negative peptides for the graph
posMean <- round(mean(df[which(df$Hemolytic == 1),1])) 
negMean <- round(mean(df[which(df$Hemolytic == 0),1]))

# Ggplots
hemo_sequences_plot <-
  ggplot(df[which(df$Hemolytic == 1),], aes(x = Length)) +
  geom_histogram(binwidth = 1,
                 center = 0,
                 fill = "#FF000099",
  ) +
  theme_ridges(font_size = 12, center_axis_labels = TRUE) +
  geom_vline(data = df,
             aes(xintercept = posMean),
             linetype = "dashed"
  ) +
  scale_x_continuous(limits = c(0, 135), breaks = seq(0, 135, 10)) +
  xlab("Sequence length") + ylab("Number of sequences") + 
  ggtitle("Hemolytic peptides") +
  scale_y_continuous(limits = c(0, 150), breaks = seq(0, 150, 50)) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

nonhemo_sequences_plot <-
  ggplot(df[which(df$Hemolytic == 0), ], aes(x = Length)) +
  geom_histogram(binwidth = 1,
                 center = 0,
                 fill = "#0000FF99") +
  theme_ridges(font_size = 12, center_axis_labels = TRUE) +
  geom_vline(data = df,
             aes(xintercept = negMean),
             linetype = "dashed") +
  scale_x_continuous(
    limits = c(0, 135),
    breaks = seq(0, 135, 10),
    minor_breaks = seq(5, 135, 5)
  ) +
  xlab("Sequence length") + ylab("Number of sequences") +
  ggtitle("Non-hemolytic peptides") +
  scale_y_continuous(limits = c(0, 150), breaks = seq(0, 150, 50)) +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))

# "publication ready" figures
ggarrange(hemo_sequences_plot, nonhemo_sequences_plot, 
          labels = c("A", "B",ncol = 2, nrow = 1), 
          legend = "right", common.legend = TRUE)
```

Some more information about length distribution

```{r}
cat("Summary of sequence length distribution of hemolytic peps\n")
summary(df[which(df$Hemolytic == 1),1])
cat("Summary of sequence length distribution of non-hemolytic peps\n")
summary(df[which(df$Hemolytic == 0),1])
```

We also analyse the amino acid compositon by plotting per amino acid the
AAC for the positive, negative and random peptides

```{r}
# Load in our dataset with the AAC and DC
datasetAAC <- read.csv("/home/jeppe/Dropbox/internship_TNO/Datasets/hemoDatabases/datasetPeptideCompositionFeatures.csv")

# Easier to select correct columns
AADict <- c("A", "R", "N", "D", "C", "E", "Q", "G", "H", 
        "I", "L", "K", "M", "F", "P", "S", "T", "W", "Y", "V")
# Take the AAC for the positives and the negatives
datasetAACpos <- datasetAAC[which(datasetAAC$Hemolytic == 1), AADict]
datasetAACneg <- datasetAAC[which(datasetAAC$Hemolytic == 0), AADict]

# Take the mean and sd for each amino acid in the AAC, positives and negatives
meanAACpos <- unlist(lapply(1:ncol(datasetAACpos), function(x) mean(datasetAACpos[,x])))
sdAACpos <- unlist(lapply(1:ncol(datasetAACpos), function(x) sd(datasetAACpos[,x])))

meanAACneg <- unlist(lapply(1:ncol(datasetAACpos), function(x) mean(datasetAACneg[,x])))
sdAACneg <- unlist(lapply(1:ncol(datasetAACpos), function(x) sd(datasetAACneg[,x])))

# Load random sequences I just downloaded from SwissProt
randomSequences4to40 <- unique(readLines("/home/jeppe/Dropbox/internship_TNO/Datasets/SwissProt/uniprot-length[4+TO+40]+reviewed+yes.fasta")[c(FALSE, TRUE)])
  # Remove non 20 canonical amino acids, just remove every peptide with weird amino acids
randomSequences4to40 <- randomSequences4to40[!grepl(paste0(c("B", "X", "J", "O", "Z", "U"),collapse = "|"), randomSequences4to40)] 

# Get the AAC, make df, transpose, sort columns according to AADict
datasetAACran <- t(as.data.frame(lapply(randomSequences4to40, extractAAC)))[, AADict]
# Get mean and sd per amino acid 
meandatasetAACran <- unlist(lapply(1:ncol(datasetAACran), function(x) mean(datasetAACran[,x])))
sddatasetAACran <- unlist(lapply(1:ncol(datasetAACran), function(x) sd(datasetAACran[,x])))

df_2 <- rbind(
  data.frame(AADict, meanAACpos, sdAACpos, "Hemolytic peptides", fix.empty.names = FALSE), 
  data.frame(AADict, meanAACneg, sdAACneg, "Non-hemolytic peptides", fix.empty.names = FALSE), 
  data.frame(AADict, meandatasetAACran, sddatasetAACran, "Swiss-Prot peptides", fix.empty.names = FALSE))
colnames(df_2) <- c("AminoAcids", "Frequency", "SD", "Group")
```


Make a plot out of it. I dont know how relevant the SD is here

```{r}
aacPlotSD <- ggplot(data=df_2, aes(x=AminoAcids, y=Frequency, fill= Group)) +
  geom_bar(stat="identity", position=position_dodge()) + scale_fill_manual(values=c('#e41a1c','#386cb0','#4daf4a')) + theme_ridges(font_size = 12,center_axis_labels = TRUE) + xlab("Amino acid") + ylab("Mean frequency") + ggtitle("Amino acid frequency for hemolytic, non-hemolytic and Swiss-Prot peptides") + theme(plot.title = element_text(hjust = 0.5)) +
  geom_errorbar(aes(ymin=Frequency-SD, ymax=Frequency+SD), width=.2,
                 position=position_dodge(.9)) 
aacPlotSD
```

One without SD

```{r}
aacPlot <- ggplot(data=df_2, aes(x=AminoAcids, y=Frequency)) +
  geom_bar(aes(fill = Group), stat="identity", position=position_dodge()) + scale_fill_manual(values=c('#e41a1c','#386cb0','#4daf4a')) + theme_ridges(font_size = 12,center_axis_labels = TRUE) + xlab("Amino acid") + ylab("Mean frequency") + ggtitle("Amino acid frequency for hemolytic, non-hemolytic and Swiss-Prot peptides") + theme(plot.title = element_text(hjust = 0.5))
aacPlot
```

Of course it is nice to know if a t-test or wilcoxon test supplies us with more
statistical insights here on differences between groups

```{r}
# Add group to datasets with AAC info
datasetAACpos$Group <- "Hemolytic peptides"
datasetAACneg$Group <- "Non-hemolytic peptides"
datasetAACran <- data.frame(datasetAACran)
datasetAACran$Group <- "Swiss-Prot peptides"
# Combine into one df
dfAACCombined <- rbind(datasetAACpos, datasetAACneg, datasetAACran)

# Is it normally distributed? (no), but also not symtrically distributed on both sides of the median...
shapiro.test(sample(dfAACCombined$A, 5000))
shapiro.test(sample(dfAACCombined$K, 5000))

# So to do test or wilcoxon? I dont know
# ttest
tTestMultiple <- function(col, dta){
  dta %>%
    rstatix::t_test(as.formula(paste0(col," ~ Group")),
                             p.adjust.method = "bonferroni")
}
tTestComparison <- lapply(AADict ,tTestMultiple, dta = dfAACCombined)
names(tTestComparison) <- AADict

# Wilcoxon
wilcoxonMultiple <- function(col, dta){
  dta %>%
    rstatix::wilcox_test(as.formula(paste0(col," ~ Group")),
                             p.adjust.method = "bonferroni")
}
wilcoxonComparison <- lapply(AADict ,wilcoxonMultiple, dta = dfAACCombined)
names(wilcoxonComparison) <- AADict

tTestComparison[[1]] # this makes sense
wilcoxonComparison[[1]] # This does not
```

We can do the same for the SSC. We will compare the SSC of the pos, neg and porter5 proteins as "normal"/comparison

```{r}
# Get positive sequences
expPos <- as.vector(readLines("/home/jeppe/Dropbox/internship_TNO/Datasets/hemoDatabases/fastas/expHemo.fasta")[c(FALSE, TRUE)])
# Get negative sequences
expNeg <- as.vector(readLines("/home/jeppe/Dropbox/internship_TNO/Datasets/hemoDatabases/fastas/expNonHemo.fasta")[c(FALSE, TRUE)])

# Get SSC
SS3 <- read.csv("/home/jeppe/Dropbox/internship_TNO/Datasets/SS3/ss3Df.csv")
SS3 <- SS3[,c("X","Coil", "Helix", "Sheet")]

# Match SSCs for pos and neg
SS3_pos <- SS3[SS3$X %in% expPos,]
SS3_pos$X <- NULL
SS3_neg <- SS3[SS3$X %in% expNeg,]
SS3_neg$X <- NULL

# Calculate mean and SD
meanSSCpos <- unlist(lapply(1:ncol(SS3_pos), function(x) mean(SS3_pos[,x])))
sdSSCpos <- unlist(lapply(1:ncol(SS3_pos), function(x) sd(SS3_pos[,x])))

meanSSCneg <- unlist(lapply(1:ncol(SS3_neg), function(x) mean(SS3_neg[,x])))
sdSSCneg <- unlist(lapply(1:ncol(SS3_neg), function(x) sd(SS3_neg[,x])))

# Import Porter5 datasets
Porter5 <- c(readLines("/home/jeppe/Dropbox/internship_TNO/Datasets/SS3/Porter5_SS/test.set"),
             readLines("/home/jeppe/Dropbox/internship_TNO/Datasets/SS3/Porter5_SS/train.set"))
# We only need 3. and 4. as information
sequencePorter5 <- Porter5[c(FALSE, FALSE, TRUE, FALSE, FALSE)]
structurePorter5 <- Porter5[c(FALSE, FALSE, FALSE, TRUE, FALSE)]

# Remove unnatural amino acids, regex so I can remove information from both vectors
AADict <- c("A", "R", "N", "D", "C", "E", "Q", "G", "H", 
        "I", "L", "K", "M", "F", "P", "S", "T", "W", "Y", "V")
regex <- paste(c("[^", paste(AADict, collapse = "|"), "]"), collapse = "")
# Which sequences do not have amino acids other than the above?
keep <- !(grepl(regex, sequencePorter5))
# We keep those
sequencePorter5 <- sequencePorter5[keep]
structurePorter5 <- structurePorter5[keep]

# Set to three class structures
structurePorter5 <- gsub(pattern = "G|H|I", replacement = "H", x = structurePorter5)
structurePorter5 <- gsub(pattern = "B|E", replacement = "E", x = structurePorter5)
structurePorter5 <- gsub(pattern = "[T|S|L|\\.]", replacement = "C", x = structurePorter5)

# Function to get SSC from porter5 data
extractSSC <- function(SS) {
  SSDict <- c("C", "H", "E")
  SSC <- as.vector(summary(factor(strsplit(SS, split = "")[[1]], levels = SSDict), 
        maxsum = 4)/nchar(SS))
}

# Get SSCs from porter5 data
SSCPorter5 <- t(as.data.frame(lapply(structurePorter5, extractSSC)))
# Get mean and SD
meanSSCPorter5 <- unlist(lapply(1:ncol(SSCPorter5), function(x) mean(SSCPorter5[,x])))
sdSSCPorter5 <- unlist(lapply(1:ncol(SSCPorter5), function(x) sd(SSCPorter5[,x])))


# Create DF for ggplot
SSDict <- c("Coil", "Helix", "Sheet")
df_3 <- rbind(
  data.frame(SSDict, meanSSCpos, sdSSCpos, "Hemolytic peptides", fix.empty.names = FALSE), 
  data.frame(SSDict, meanSSCneg, sdSSCneg, "Non-hemolytic peptides", fix.empty.names = FALSE), 
  data.frame(SSDict, meanSSCPorter5, sdSSCPorter5, "Porter5 proteins", fix.empty.names = FALSE))
colnames(df_3) <- c("SSclasses", "Frequency", "SD", "Group")
```

Now we plot the SSC of pos, neg and porter

```{r}
sscPlotSD <- ggplot(data=df_3, aes(x=SSclasses, y=Frequency, fill= Group)) +
  geom_bar(stat="identity", position=position_dodge()) + scale_fill_manual(values=c('#e41a1c','#386cb0','#4daf4a')) + theme_ridges(font_size = 12,center_axis_labels = TRUE) + xlab("Secondary structure") + ylab("Mean frequency") + ggtitle("Secondary structure frequency for hemolytic, non-hemolytic and Porter5 proteins") + theme(plot.title = element_text(hjust = 0.5)) +
  geom_errorbar(aes(ymin=Frequency-SD, ymax=Frequency+SD), width=.2,
                 position=position_dodge(.9)) 
sscPlotSD
```

One without SD

```{r}
sscPlot <- ggplot(data=df_3, aes(x=SSclasses, y=Frequency, fill= Group)) +
  geom_bar(stat="identity", position=position_dodge()) + scale_fill_manual(values=c('#e41a1c','#386cb0','#4daf4a')) + theme_ridges(font_size = 12,center_axis_labels = TRUE) + xlab("Secondary structure") + ylab("Mean frequency") + ggtitle("Secondary structure frequency for hemolytic, \nnon-hemolytic and Porter5 proteins") + theme(plot.title = element_text(hjust = 0.5))
sscPlot
```

Do the SSC groupd differ from each other? Again wilcoxon and ttest

```{r}
SS3_pos$Group <- "Hemolytic peptides"
SS3_neg$Group <- "Non-hemolytic peptides"
SSCPorter5 <- data.frame(SSCPorter5)
SSCPorter5$Group <- "Swiss-Prot peptides"
colnames(SSCPorter5) <- colnames(SS3_neg)
dfSSCCombined <- rbind(SS3_pos, SS3_neg, SSCPorter5)

# Is it normally distributed? (no), but also not symtrically distributed on both sides of the median...
shapiro.test(sample(dfSSCCombined$Coil, 5000))
shapiro.test(sample(dfSSCCombined$Sheet, 5000))

# So to do test or wilcoxon? I dont know
tTestComparisonSSC <- lapply(c("Coil", "Helix", "Sheet") ,tTestMultiple, dta = dfSSCCombined)
names(tTestComparison) <- c("Coil", "Helix", "Sheet")

wilcoxonComparisonSSC <- lapply(c("Coil", "Helix", "Sheet") ,wilcoxonMultiple, dta = dfSSCCombined)
names(wilcoxonComparison) <- c("Coil", "Helix", "Sheet")

tTestComparisonSSC[[1]] # this makes sense
wilcoxonComparisonSSC[[1]] # This does not
```

Also, lets see if we can visualise the peptides based on AAC, DC, and SSC

```{r}
set.seed(123)
# Store class
Y <- as.character(datasetAAC$Hemolytic)
sequences <- datasetAAC$Sequence
# Split data
# DC
datasetDC <- as.matrix(datasetAAC[,31:430])
# AAC + DC + SSC
datasetAAC_DC_SSC <- as.matrix(datasetAAC[,8:430])
# AAC
datasetAAC <- as.matrix(datasetAAC[,AADict])

datasetList <- lapply(list(datasetAAC, datasetDC, datasetAAC_DC_SSC), scale)
# Run Umap on AAC, DC and AAC + DC + SSC
umapList <- lapply(datasetList, umap)
# Run TSNE on AAC, DC and AAC + DC + SSC
tsneList <- lapply(datasetList, Rtsne, check_duplicates = FALSE)

```

Now that we have the UMAP and TSNE dimensionality reduction results we can use 
the first two dims of each to create dot plots.

```{r}
# Plot Umap for AAC, DC and AAC + DC + SSC
umapPlotList <- lapply(umapList, function(x) {
  # Store in df
  df <- data.frame(x = x[, 1], 
                                  y = x[,2], hemo = Y,
                  stringsAsFactors = FALSE)
# Plot UMAP of peptides, colour by hemolytic class
ggplot(df, aes(x = x, y = y, color = hemo)) +
  geom_point(size = 1.3) + theme_void() + scale_color_manual(values = c('#386cb0', '#e41a1c'))
  }
)

# Plot TSNE on AAC, DC and AAC + DC + SSC
tsnePlotList <- lapply(tsneList, function(x) {
  # Store in df
  df <- data.frame(x = x[["Y"]][, 1], 
                                  y = x[["Y"]][,2], hemo = Y,
                  stringsAsFactors = FALSE)
# Plot TSNE of peptides, colour by hemolytic class
ggplot(df, aes(x = x, y = y, color = hemo)) +
  geom_point(size = 1.3) + theme_void() + scale_color_manual(values = c('#386cb0', '#e41a1c'))
  }
)

# plot the plots

# UMAP
for (i in 1:3){
  plot(umapPlotList[[i]])
}
# TSNE
for (i in 1:3){
  plot(tsnePlotList[[i]])
}
```

Lets continue with the AAC, DC, SSC of the TSNEvisualisation to see where in 
TSNE different information on the peptide composition clusters

```{r}
# Store tsne of aac dc and ssc
x <- tsneList[[3]]
# create df with tsne 2 dims, hemo info
df <- data.frame(x = x[["Y"]][, 1], y = x[["Y"]][,2], hemo = ifelse(Y == 1, "Hemolytic", "Non-hemolytic"),
                  stringsAsFactors = FALSE)
# add AAC, DC, SSC info
df <- cbind(df, datasetAAC_DC_SSC)

# plot AAC and SSC info on the TSNE, put it in a function so i dont have to
# copy paste all the time
plotPlots <- function() {
plot(ggplot(df, aes(x = x, y = y, color = hemo)) +
  geom_point(size = 1.3) + theme_void()  + 
  scale_color_manual(values = c('#e41a1c', '#386cb0'), name = "Class") + 
  ggtitle("Hemolytic and non-hemolytic peptides") +
  theme(plot.title = element_text(hjust = 0.5)))

plot(ggplot(df, aes(x = x, y = y, color = Coil)) +
  geom_point(size = 1.3) + theme_void() +
  scale_color_viridis(option = "D")+ 
  ggtitle("Coil") +
  theme(plot.title = element_text(hjust = 0.5)))

plot(ggplot(df, aes(x = x, y = y, color = Helix)) +
  geom_point(size = 1.3) + theme_void() +
  scale_color_viridis(option = "D")+ 
  ggtitle("Helix") +
  theme(plot.title = element_text(hjust = 0.5)))

plot(ggplot(df, aes(x = x, y = y, color = Sheet)) +
  geom_point(size = 1.3) + theme_void() +
  scale_color_viridis(option = "D")+ 
  ggtitle("Sheet") +
  theme(plot.title = element_text(hjust = 0.5)))


for (aa in AADict){
  a <- ggplot(df, aes_string(x = "x", y = "y", color = aa)) +
  geom_point(size = 1.3) + theme_void() +
  scale_color_viridis(option = "D") + 
  ggtitle(paste0("Amino acid: ", aa)) +
  theme(plot.title = element_text(hjust = 0.5))
  plot(a)
}
}

plotPlots()
```

Lets put the Hemolytic class, Coil, Sheet, C, L and W in a nice plot for my
thesis

```{r}
a1 <- ggplot(df, aes(x = x, y = y, color = hemo)) +
  geom_point(size = 1.3) + theme_void()  + 
  scale_color_manual(values = c('#e41a1c', '#386cb0'), name = "Class") + 
  ggtitle("Hemolytic and non-hemolytic peptides") +
  theme(plot.title = element_text(hjust = 0.5))
a2 <- ggplot(df, aes(x = x, y = y, color = Coil)) +
  geom_point(size = 1.3) + theme_void() +
  scale_color_viridis(option = "D")+ 
  ggtitle("Coil") +
  theme(plot.title = element_text(hjust = 0.5))
a3 <- ggplot(df, aes(x = x, y = y, color = Helix)) +
  geom_point(size = 1.3) + theme_void() +
  scale_color_viridis(option = "D")+ 
  ggtitle("Helix") +
  theme(plot.title = element_text(hjust = 0.5))
a4 <- ggplot(df, aes(x = x, y = y, color = Sheet)) +
  geom_point(size = 1.3) + theme_void() +
  scale_color_viridis(option = "D")+ 
  ggtitle("Sheet") +
  theme(plot.title = element_text(hjust = 0.5))
a5 <- ggplot(df, aes(x = x, y = y, color = L)) +
  geom_point(size = 1.3) + theme_void() +
  scale_color_viridis(option = "D")+ 
  ggtitle("Leucine") +
  theme(plot.title = element_text(hjust = 0.5))
a6 <- ggplot(df, aes(x = x, y = y, color = K)) +
  geom_point(size = 1.3) + theme_void() +
  scale_color_viridis(option = "D")+ 
  ggtitle("Lysine") +
  theme(plot.title = element_text(hjust = 0.5))


ggarrange(a1,a2,a3,a4,a5,a6, labels = c("A", "B", "C", "D", "E", "F"), common.legend = F, legend = "right", ncol = 2, nrow = 3)
```



Maybe we can cluster the peptides? 
Lets first scale and center the data

```{r}
scaledDatasetAAC_DC_SSC <- scale(datasetAAC_DC_SSC)
```


Lets try Kmeans clustering


```{r}
set.seed(2)
# Kmeans clustering
# Lets try for the AAC DC SSC data, because that is what worked best for the RF
km5 <- eclust(datasetList[[3]], FUNcluster = "kmeans", iter.max = 25, k = 5)
km10 <- eclust(datasetList[[3]], FUNcluster = "kmeans", iter.max = 25, k = 10)
km15 <- eclust(datasetList[[3]], FUNcluster = "kmeans", iter.max = 25, k = 15)
km20 <- eclust(datasetList[[3]], FUNcluster = "kmeans", iter.max = 25, k = 20)
km25 <- eclust(datasetList[[3]], FUNcluster = "kmeans", iter.max = 25, k = 25)
km30 <- eclust(datasetList[[3]], FUNcluster = "kmeans", iter.max = 25, k = 30)
# TSNE hemolytic
ggplot(df, aes(x = x, y = y, color = hemo)) +
  geom_point(size = 1.3) + theme_void()  + scale_color_manual(values = c('#386cb0', '#e41a1c'))

# KM clusters tsne plot
ggplot(df, aes(x = x, y = y, color = as.factor(km5$cluster) )) +
  geom_point(size = 1.3) + theme_void() 

ggplot(df, aes(x = x, y = y, color = as.factor(km15$cluster) )) +
  geom_point(size = 1.3) + theme_void()

ggplot(df, aes(x = x, y = y, color = as.factor(km20$cluster) )) +
  geom_point(size = 1.3) + theme_void()

ggplot(df, aes(x = x, y = y, color = as.factor(km25$cluster) )) +
  geom_point(size = 1.3) + theme_void()

ggplot(df, aes(x = x, y = y, color = as.factor(km30$cluster) )) +
  geom_point(size = 1.3) + theme_void()

plotPlots()

# Km with k = 10 and 15 does not even look that stupid, especially since cluster 3
# seems to cluster on a high fraction of amino acid K
# Seems to catch some meaning full clusters? K = 20 gets hard to see with these
# colours.
```

We can try some methods to determine which K to use, elbow, silouette and gap

```{r}
# Elbow
elbow <- fviz_nbclust(datasetAAC_DC_SSC, kmeans, method = "wss", k.max = 40, iter.max = 30)
elbow
# SIlhouetten
sil <- fviz_nbclust(datasetAAC_DC_SSC, kmeans, method = "silhouette", k.max = 40, iter.max = 30)
sil
# Gap
gap <- fviz_nbclust(datasetAAC_DC_SSC, kmeans, method = "gap_stat", k.max = 40, iter.max = 30)
```

They all give k = 2... I dont think I have the time and means to find biolgical
meaningfull clusters from this
Lets store some for my report
```{r}
# get some good colours
# R colours brewer has a nice palette. I just combine all and then sample an n
library(RColorBrewer)
# Store colours
qualCols = brewer.pal.info[brewer.pal.info$category == 'qual',]
colVector = unlist(mapply(brewer.pal, qualCols$maxcolors, rownames(qualCols)))

# k5 plot
k5TSNE <- ggplot(df, aes(x = x, y = y, color = as.factor(km$km5) )) +
  geom_point(size = 1.3) + theme_void() + 
  ggtitle("Clustering with k = 5") +
  theme(plot.title = element_text(hjust = 0.5)) + labs(color = "Clusters") + 
  scale_color_manual(values = sample(colVector, 5))
k5TSNE
# k30 plot
k30TSNE <- ggplot(df, aes(x = x, y = y, color = as.factor(km$km30) )) +
  geom_point(size = 1.3) + theme_void() + 
  ggtitle("Clustering with k = 30") +
  theme(plot.title = element_text(hjust = 0.5)) + labs(color = "Clusters")+ 
  scale_color_manual(values = sample(colVector, 30))
k30TSNE

# Arrange
ggarrange(k5TSNE, k30TSNE, labels = c("A", "B"), common.legend = T, legend = "none", ncol = 2)
# Save it because I did not set seed and these are nice.
saveRDS(ggarrange(k5TSNE, k30TSNE, labels = c("A", "B"), common.legend = T, legend = "none", ncol = 2), "/home/jeppe/Dropbox/internship_TNO/Figures/Results/clustersk5k30.rds")
```


CDHIT clusters? I have 90-60 redundancy reduced and 90-60-40

```{r}
# Read cdhit results, start with 90 60
cdhit90_60 <- readLines("/home/jeppe/Dropbox/internship_TNO/Datasets/cdhit90-60/1600291656.fas.2-0.clstr.sorted")
# clusters are marked by >Cluster, find where those cluster lines start
clusters <- grep(">Cluster", cdhit90_60)
# For each cluster we do:
clustersList <- lapply(1:length(clusters), function(x){
  # If it is the last cluster in the list
  if (x == length(clusters)) {
    # Store the peptides
    vec <- cdhit90_60[clusters[x]:length(cdhit90_60)]
  }
  else{
    # STore the peptides between >cluster x and >cluster y
    y <- x + 1
  vec <- cdhit90_60[clusters[x]:clusters[y]]
  }
})

# Grep peptides per list entry, format is peptide_ and a number 1 to 4 times
# Use str_extract, remove NA's
clustersList <- lapply(1:length(clustersList), function(x){
  peps <- na.omit(str_extract(clustersList[[x]], pattern = ">peptide_[0-9]{1,4}"))
})

# Read in our fasta
fasta <- readLines("/home/jeppe/Dropbox/internship_TNO/Datasets/hemoDatabases/fastas/sequences.fasta")

# get the fasta headers
sequenceNames <- fasta[c(TRUE,FALSE)]
# and the sequences
sequencesFasta <- fasta[c(FALSE,TRUE)]

# Now the sequences in the fasta file are sorted differently then the sequences
# from the database with peptide compositions (That we used for the TSNE). 
# We need to sort the names thus like the sequences in the database with peptide 
# compositions
names(sequenceNames) <- sequencesFasta
sequenceNames <- sequenceNames[as.character(sequences)]

# Create df to add clusters to
clustersCDHIT9060 <-  as.data.frame(sequenceNames)
# Add a column for clusters, we had to remove sequences smaller then 5 in CD-HIT
# so we add missing as standard annotation
clustersCDHIT9060$Cluster <- "Missing"
# Add per peptide what cluster they belong to
for(x in 1:length(clustersList)) {
  pepsInCluster <- clustersList[[x]] # Which peptides are in this cluster?
  # For clusters, if there are more then 1 peptide in the cluster
  if(length(pepsInCluster) >= 2){
  # Add for each paptide which cluster they belong to
  clustersCDHIT9060[as.character(clustersCDHIT9060$sequenceNames) %in% pepsInCluster, 2] <- paste0("Cluster_", x)}
  else{
  clustersCDHIT9060[as.character(clustersCDHIT9060$sequenceNames) %in% pepsInCluster, 2] <- "Singleton"} 
  }

# How many clusters (excluding singletons) are there anyway?
length(unique(grep("Cluster", x = clustersCDHIT9060$Cluster, value = TRUE)))
# 487

# Show clusters on tsne
ggplot(df, aes(x = x, y = y, color = clustersCDHIT9060$Cluster)) +
  geom_point(size = 1.3) + theme_void() + theme(legend.position = "none") + 
  ggtitle("CD-Hit clusters") +
  theme(plot.title = element_text(hjust = 0.5))

plotPlots()
# Does seem to contain also some info? Red cluster, cluster on high L and high K?
# Brown glob at the bottom is also the same as the red glob at the bottom of k = 15 above
```

Same for90 60 40 

```{r}
# Read cdhit results
cdhit90_60_40 <- readLines("/home/jeppe/Dropbox/internship_TNO/Datasets/cdhit90-60-40/1600291504.fas.3-0.clstr.sorted")
# clusters are marked by >Cluster, find where those cluster lines start
clusters <- grep(">Cluster", cdhit90_60_40)
# For each cluster we do:
clustersList <- lapply(1:length(clusters), function(x){
  # If it is the last cluster in the list
  if (x == length(clusters)) {
    # Store the peptides
    vec <- cdhit90_60_40[clusters[x]:length(cdhit90_60_40)]
  }
  else{
    # STore the peptides between >cluster x and >cluster y
    y <- x + 1
  vec <- cdhit90_60_40[clusters[x]:clusters[y]]
  }
})

# Grep peptides per list entry, format is peptide_ and a number 1 to 4 times
# Use str_extract, remove NA's
clustersList <- lapply(1:length(clustersList), function(x){
  peps <- na.omit(str_extract(clustersList[[x]], pattern = ">peptide_[0-9]{1,4}"))
})

# Read in our fasta
fasta <- readLines("/home/jeppe/Dropbox/internship_TNO/Datasets/hemoDatabases/fastas/sequences.fasta")

# get the fasta headers
sequenceNames <- fasta[c(TRUE,FALSE)]
# and the sequences
sequencesFasta <- fasta[c(FALSE,TRUE)]

# Now the sequences in the fasta file are sorted differently then the sequences
# from the database with peptide compositions (That we used for the TSNE). 
# We need to sort the names thus like the sequences in the database with peptide 
# compositions
names(sequenceNames) <- sequencesFasta
sequenceNames <- sequenceNames[as.character(sequences)]

# Create df to add clusters to
clustersCDHIT906040 <-  as.data.frame(sequenceNames)
# Add a column for clusters, we had to remove sequences smaller then 5 in CD-HIT
# so we add missing as standard annotation
clustersCDHIT906040$Cluster <- "Missing"
# Add per peptide what cluster they belong to
for(x in 1:length(clustersList)) {
  pepsInCluster <- clustersList[[x]] # Which peptides are in this cluster?
  # For clusters, if there are more then 1 peptide in the cluster
  if(length(pepsInCluster) >= 2){
  # Add for each paptide which cluster they belong to
  clustersCDHIT906040[as.character(clustersCDHIT906040$sequenceNames) %in% pepsInCluster, 2] <- paste0("Cluster_", x)}
  else{
  clustersCDHIT906040[as.character(clustersCDHIT906040$sequenceNames) %in% pepsInCluster, 2] <- "Singleton"} 
  }

# How many clusters (excluding singletons) are there anyway?
length(unique(grep("Cluster", x = clustersCDHIT906040$Cluster, value = TRUE)))
# 200

# Show clusters on tsne
ggplot(df, aes(x = x, y = y, color = clustersCDHIT906040$Cluster)) +
  geom_point(size = 1.3) + theme_void() + theme(legend.position = "none") + 
  ggtitle("CD-Hit clusters") +
  theme(plot.title = element_text(hjust = 0.5))

plotPlots()
```

Store these clusters

```{r}
# Create DF
clustersDF <- clustersCDHIT9060
# Change colnames
colnames(clustersDF) <- c("fastaName", "clustersCDHIT9060")
# Store sequences as proper column
clustersDF$Sequences <- rownames(clustersDF)

# same for 90-60-40
colnames(clustersCDHIT906040) <- c("fastaName", "clustersCDHIT906040")

# join
clustersDF <- inner_join(clustersDF, clustersCDHIT906040)
# add KM clusters
clustersDF$km5 <- km5$cluster
clustersDF$km10 <- km10$cluster
clustersDF$km15 <- km15$cluster
clustersDF$km20 <- km20$cluster
clustersDF$km25 <- km25$cluster
clustersDF$km30 <- km30$cluster

write.csv(clustersDF, "/home/jeppe/Dropbox/internship_TNO/R/6_AnalysesAndFigures/clustersPeptideComposition/clustersPeptideComp.csv")
```

