```{r}
library(ComplexHeatmap)
library(uwot)
library(stringr)
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(Rtsne)
library(RColorBrewer)
library(ggthemes)
library(dplyr)
library(factoextra)
library(parallel)
library(doParallel)
library(word2vec)
library(viridis)
library(dbscan)
```

Let's take a closer look at the word2vec features we have used

```{r}
# load our models
modelSentences1gram <- read.word2vec("/home/jeppe/Dropbox/internship_TNO/R/5_Word2Vec/Word2VecModels/sentences1GramCombinedDim300Window25.bin")
modelSentences2gram <- read.word2vec("/home/jeppe/Dropbox/internship_TNO/R/5_Word2Vec/Word2VecModels/sentences2gramCombinedDim300Window25.bin")
```

## Analysis of the embeddings
3-grams of the primary structure only, I use the one from the hemolytic
and non-hemolytic peptides. Let's see if it learned something
```{r}
embeddingsPrimStruc <- as.matrix(read.word2vec("/home/jeppe/Dropbox/internship_TNO/R/5_Word2Vec/Word2VecModels/modelHemoSequence3gram300Dim.bin"))

# Create Umap so we can visually inspect how words relate to each other and guess clusters
set.seed(200)

# UMAP for 1grams
umap3GramSeqFeatures <- umap(embeddingsPrimStruc, n_threads = 4)

# Store in df for ggplot
df3GramSeqFeaturesUmap  <- data.frame(word = gsub("//.+", "", rownames(embeddingsPrimStruc)), 
                  xpos = gsub(".+//", "", rownames(embeddingsPrimStruc)), 
                  x = umap3GramSeqFeatures[, 1], y = umap3GramSeqFeatures[, 2], 
                  stringsAsFactors = FALSE)
# Plot umap of 3 gram
ggplot(df3GramSeqFeaturesUmap, aes(x = x, y = y)) +
  geom_point() + theme_void()

# TSNE for 3GramSeqs
tsne3GramSeqFeatures <- Rtsne(embeddingsPrimStruc, pca_center = F, normalize = F, initial_dims = 10)
# Store in df for ggplot
df3GramSeqFeaturesTsne  <- data.frame(word = gsub("//.+", "", rownames(embeddingsPrimStruc)), 
                  xpos = gsub(".+//", "", rownames(embeddingsPrimStruc)), 
                  x = tsne3GramSeqFeatures[["Y"]][, 1], y = tsne3GramSeqFeatures[["Y"]][, 2], 
                  stringsAsFactors = FALSE)
# Plot
ggplot(df3GramSeqFeaturesTsne, aes(x = x, y = y, label = word)) +
  geom_point() + theme_void()
```

Lets see if these word2vec representations contain biological information

```{r}
# Dictionary for the hydrophobicity
hydrophobicity <-
  c(
    "A" = 0.62,
    "R" = -2.53 ,
    "N" = -0.78,
    "D" = -0.90,
    "C" = 0.29,
    "E" = -0.74,
    "Q" = -0.85,
    "G" = 0.48,
    "H" = -0.40,
    "I" = 1.38,
    "L" = 1.06,
    "K" = -1.5,
    "M" = 0.64,
    "F" = 1.19,
    "P" = 0.12,
    "S" = -0.18,
    "T" = -0.05,
    "W" = 0.81,
    "Y" = 0.26,
    "V" = 1.08
  ) # JACS, 1962, 84: 4240-4246. (C. Tanford).
hydrophilicy <-
  c(
    "A" = -0.5,
    "R" = 3,
    "N" = 0.2,
    "D" = 3,
    "C" = -1,
    "E" = 3,
    "Q" = 0.2,
    "G" = 0,
    "H" = -0.5,
    "I" = -1.8,
    "L" = -1.8,
    "K" = 3,
    "M" = -1.3,
    "F" = -2.5,
    "P" = 0,
    "S" = 0.3,
    "T" = -0.4,
    "W" = -3.4,
    "Y" = -2.3,
    "V" = -1.5
  ) # PNAS, 1981, 78:3824-3828 (T.P.Hopp & K.R.Woods)..

charge <-
  c(
    "A" = 0,
    "R" = 1,
    "N" = 0,
    "D" = -1,
    "C" = 0,
    "E" = -1,
    "Q" = -0,
    "G" = 0,
    "H" = 1,
    "I" = 0,
    "L" = 0,
    "K" = 1,
    "M" = 0,
    "F" = 0,
    "P" = -0,
    "S" = -0,
    "T" = 0,
    "W" = 0 ,
    "Y" = 0,
    "V" = 0
  )
mass <- # side chain mass!
  c(
    "A" = 15,
    "R" = 101,
    "N" = 58,
    "D" = 59,
    "C" = 47,
    "E" = 73,
    "Q" = 72,
    "G" = 1,
    "H" = 82,
    "I" = 57,
    "L" = 57,
    "K" = 73,
    "M" = 75,
    "F" = 91,
    "P" = 42,
    "S" = 31,
    "T" = 45,
    "W" = 130 ,
    "Y" = 107,
    "V" = 43
  )
# Take sum of the above chemical properties for amino acids in a "word"
chemicalPropertiesembPrimStruc <- t(as.data.frame(lapply(rownames(embeddingsPrimStruc), function(x){
  aminoAcids <- str_split(x, simplify = TRUE, pattern = "")
  return(c(sum(hydrophobicity[aminoAcids]), sum(hydrophilicy[aminoAcids]), sum(charge[aminoAcids]), sum(mass[aminoAcids])))
})))

# add row and colnames 
rownames(chemicalPropertiesembPrimStruc) <- rownames(embeddingsPrimStruc)
colnames(chemicalPropertiesembPrimStruc) <- c("Hydrophobicity", "Hydrophilicy","Charge", "Mass")

# Umap
# Plot
p1 <- ggplot(df3GramSeqFeaturesUmap, aes(x = x, y = y, color = chemicalPropertiesembPrimStruc[,1])) +
  geom_point() + theme_void() +
  scale_color_viridis(option = "D", labels=c("Minimum", "", "", "", "Maximum", "Maximum")) + ggtitle("Hydrophobicity") +
  theme(plot.title = element_text(hjust = 0.5))
p1[["labels"]][["colour"]] <- ""
# Plot
p2 <- ggplot(df3GramSeqFeaturesUmap, aes(x = x, y = y, color = chemicalPropertiesembPrimStruc[,2])) +
  geom_point() + theme_void() +
  scale_color_viridis(option = "D")  + ggtitle("Hydrophilicy") +
  theme(plot.title = element_text(hjust = 0.5))

# Plot
p3 <- ggplot(df3GramSeqFeaturesUmap, aes(x = x, y = y, color = chemicalPropertiesembPrimStruc[,3])) +
  geom_point() + theme_void() +
  scale_color_viridis(option = "D")  + ggtitle("Charge") +
  theme(plot.title = element_text(hjust = 0.5))

p4 <- ggplot(df3GramSeqFeaturesUmap, aes(x = x, y = y, color = chemicalPropertiesembPrimStruc[,4])) +
  geom_point() + theme_void() +
  scale_color_viridis(option = "D")  + ggtitle("Mass") +
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(p1, p2, p3, p4, labels = c("A", "B", "C", "D"), common.legend = TRUE, legend = "right")
```

Lets analyse the 1-gram embeddings of the primary sequence and sec structure
```{r}
# Store our embeddings per "word" 
embedding1Gram <- as.matrix(modelSentences1gram)

# </s> is a stop signal, which is basically the end of the sentence. 
# You would need this in a text
# I see no reason to keep this for a peptide sequence, so: remove
embedding1Gram <- embedding1Gram[-which(rownames(embedding1Gram) == "</s>"),]
# Add space between amino acid and structure
rownames(embedding1Gram) <- gsub("(^.)", "\\1 ", rownames(embedding1Gram))
# Create Umap so we can visually inspect how words relate to each other and guess clusters
set.seed(2)

# UMAP for 1grams
umap1GramFeatures <- umap(embedding1Gram, n_neighbors = 15, n_threads = 2, spread = 10)

# Store in df for ggplot
df1GramFeaturesUmap  <- data.frame(word = gsub("(^.)", "\\1 ", rownames(embedding1Gram)), 
                  xpos = gsub(".+//", "", rownames(embedding1Gram)), 
                  x = umap1GramFeatures[, 1], y = umap1GramFeatures[, 2], 
                  stringsAsFactors = FALSE)
# Plot umap of 1 gram
ggplot(df1GramFeaturesUmap, aes(x = x, y = y, label = word)) +
  geom_point() + geom_text_repel() + theme_void()

# TSNE for 1grams
tsne1GramFeatures <- Rtsne(embedding1Gram, perplexity = 10)
# Store in df for ggplot
df1GramFeaturesTsne  <- data.frame(word = gsub("(^.)", "\\1 ", rownames(embedding1Gram)), 
                  xpos = gsub(".+//", "", rownames(embedding1Gram)), 
                  x = tsne1GramFeatures[["Y"]][, 1], y = tsne1GramFeatures[["Y"]][, 2], 
                  stringsAsFactors = FALSE)
# Plot
ggplot(df1GramFeaturesTsne, aes(x = x, y = y, label = word)) +
  geom_point() + geom_text_repel() + theme_void()
```

Lets see where the chemical properties cluster, not that interesting for 1 grams
since there is not much points to visualy analayse and to see where stuff clusters

```{r}
# Take mean of the above chemical properties for amino acids in a "word"
# for a 1gram not that interesting
chemicalPropertiesemb1Grams <- t(as.data.frame(lapply(rownames(embedding1Gram), function(x){
  aminoAcids <- str_split(x, simplify = TRUE, pattern = "")
  return(c(mean(hydrophobicity[aminoAcids[1]]),mean(hydrophilicy[aminoAcids[1]]), mean(charge[aminoAcids[1]]), mean(mass[aminoAcids[1]])))
})))
# add row and col naemes
rownames(chemicalPropertiesemb1Grams) <- rownames(embedding1Gram)
colnames(chemicalPropertiesemb1Grams) <- c("Hydrophobicity","Hydrophilicy", "Charge", "Mass")

# Plot
pp1 <- ggplot(df1GramFeaturesTsne, aes(x = x, y = y, label = word, color = chemicalPropertiesemb1Grams[,1])) +
  geom_point() + theme_void() +
  scale_color_viridis(option = "D", labels=c("Minimum", "Minimum", "", "", "Maximum"))  + ggtitle("Hydrophobicity") +
  theme(plot.title = element_text(hjust = 0.5))
pp1[["labels"]][["colour"]] <- ""

# Plot
pp2 <-ggplot(df1GramFeaturesTsne, aes(x = x, y = y, label = word,color = chemicalPropertiesemb1Grams[,2])) +
  geom_point() + theme_void() +
  scale_color_viridis(option = "D")   + ggtitle("Hydrophilicy") +
  theme(plot.title = element_text(hjust = 0.5))
# Plot
pp3 <-ggplot(df1GramFeaturesTsne, aes(x = x, y = y, label = word,color = chemicalPropertiesemb1Grams[,3])) +
  geom_point() + theme_void() +
  scale_color_viridis(option = "D")  + ggtitle("Charge") +
  theme(plot.title = element_text(hjust = 0.5))
# Plot
pp4 <-ggplot(df1GramFeaturesTsne, aes(x = x, y = y, label = word,color = chemicalPropertiesemb1Grams[,4])) +
  geom_point() + theme_void() +
  scale_color_viridis(option = "D") + ggtitle("Mass") +
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(pp1,pp2, pp3, pp4, labels = c("A", "B", "C", "D"), common.legend = TRUE, legend = "right")
```

Let's add some metadata so we can maybe see some extra info

```{r}
# Add information about the 1-grams, simply what "c;ass" the amno acid belongs to
df1GramFeaturesUmap$'Amino acid properties' <- "Hydrophobic"
df1GramFeaturesUmap[grepl("S|T", x = df1GramFeaturesUmap$word),5] <- "Polar (positive)"
df1GramFeaturesUmap[grepl("N|Q", x = df1GramFeaturesUmap$word),5] <- "Polar (negative)"
df1GramFeaturesUmap[grepl("C|G|P", x = df1GramFeaturesUmap$word),5] <- "Special-cases"
df1GramFeaturesUmap[grepl("R|H|K", x = df1GramFeaturesUmap$word),5] <- "Charge (positive)"
df1GramFeaturesUmap[grepl("D|E", x = df1GramFeaturesUmap$word),5] <- "Charge (negative)"

# Plot
ggplot(df1GramFeaturesUmap, aes(x = x, y = y, label = word, color = df1GramFeaturesUmap[,5])) + theme_void() + scale_color_manual(values = c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#999999','#a65628','#f781bf','#999999')) +
  labs(color = "Amino acid properties") + theme(legend.text=element_text(size=11)) +  coord_cartesian(clip = "off")+ 
  geom_text_repel(segment.alpha = 0) + xlim(c(-40, 35))  + ylim(c(-40, 35))+ ggtitle("UMAP of the numerical features based on 1 grams") +
  theme(plot.title = element_text(hjust = 0.5))
```


Three clear clusters of helix, sheet and coil, k = 3? Let's see what clustering 
does.

```{r}
set.seed(1)
# Kmeans clustering
km1GramFeatures <- eclust(embedding1Gram, "kmeans", iter.max = 25)
fviz_gap_stat(km1GramFeatures$gap_stat)
fviz_silhouette(km1GramFeatures)

# Hierarchical clustering
hc1GramFeatures <- eclust(embedding1Gram, "hclust")
fviz_dend(hc1GramFeatures, rect = TRUE)

# Scatter plots of both clustering methods
fviz_cluster(km1GramFeatures)
fviz_cluster(hc1GramFeatures)
```

We can visualise this in a nice heat map, also good way to see that these numerical
vectors per words have overlap and differences which you can cluster/classify with

```{r}
set.seed(1)
# Heatmap, dont cluster columns and cluster rows to 3
Heatmap(embedding1Gram, cluster_columns = FALSE, row_names_gp = gpar(fontsize = 8), show_column_dend = FALSE, show_row_dend = TRUE, row_km = 3, name = "1 Gram features\n", heatmap_legend_param = list(labels = c("Minimum", "", "", "", "Maximum")), column_title = "Heatmap of all numerical features based on 1 grams")
```

We can zoom in on three amino acids and their three classes

```{r}
# Order the data so rows are organised in heatmap
order <- str_sort(grep("G .....?|L .....?|K .....?", rownames(embedding1Gram), value = TRUE))
order <- sort(rownames(embedding1Gram[grepl("G .....?|L .....?|K .....?", rownames(embedding1Gram)),]))
# Heatmap
Heatmap(embedding1Gram[grepl("G .....?|L .....?|K .....?", rownames(embedding1Gram)),], cluster_rows = F, cluster_columns = FALSE, row_names_gp = gpar(fontsize = 11), row_order = order, row_split = c("Glycine","Leucine","Lysine","Leucine","Lysine","Leucine","Glycine","Lysine","Glycine"), name = "1 Gram features\n", heatmap_legend_param = list(labels = c("Minimum", "", "", "", "Maximum")), column_title = "Heatmap of the numerical features based on the 1 grams of glycine, \nleucine and lysine and secondary structures")
```

Lets do the same for the 2 grams, these are 3600 possible "word" representations
instead of 60

```{r}
# Load matrix with words and their numerical vectors
embedding2Gram <- as.matrix(modelSentences2gram)

# </s> is a stop signal, which is basically the end of the sentence. 
# You would need this in a text
# I see no reason to keep this for a peptide sequence, so: remove
embedding2Gram <- embedding2Gram[-which(rownames(embedding2Gram) == "</s>"),]

set.seed(2)
# Run Umap
umap2GramFeatures <- umap(embedding2Gram, n_threads = 5)

# Get in DF
df2GramFeatureUmap  <- data.frame(word = gsub("//.+", "", rownames(embedding2Gram)), 
                  xpos = gsub(".+//", "", rownames(embedding2Gram)), 
                  x = umap2GramFeatures[, 1], y = umap2GramFeatures[, 2], 
                  stringsAsFactors = FALSE)

# Run TSNE
tsne2GramFeatures <- Rtsne(embedding2Gram)

# Get in DF
df2GramFeatureTsne  <- data.frame(word = gsub("//.+", "", rownames(embedding2Gram)), 
                  xpos = gsub(".+//", "", rownames(embedding2Gram)), 
                  x = tsne2GramFeatures[["Y"]][, 1], y = tsne2GramFeatures[["Y"]][, 2], 
                  stringsAsFactors = FALSE)

# UMAP
ggplot(df2GramFeatureUmap, aes(x = x, y = y)) +
  geom_point()
# TSNE
ggplot(df2GramFeatureTsne, aes(x = x, y = y)) +
  geom_point()
```

Lets add groups. I just use the 9 possible combinations of two times three classes
as groups. Seemed the most obvious one to me. 

```{r}
df2GramFeatureUmap$Config <- "SheetSheet"
df2GramFeatureUmap[grepl(".sheet.coil", x = df2GramFeatureUmap$word),5] <- "SheetCoil"
df2GramFeatureUmap[grepl(".sheet.helix", x = df2GramFeatureUmap$word),5] <- "SheetHelix"
df2GramFeatureUmap[grepl(".coil.coil", x = df2GramFeatureUmap$word),5] <- "CoilCoil"
df2GramFeatureUmap[grepl(".coil.helix", x = df2GramFeatureUmap$word),5] <- "CoilHelix"
df2GramFeatureUmap[grepl(".coil.sheet", x = df2GramFeatureUmap$word),5] <- "CoilSheet"
df2GramFeatureUmap[grepl(".helix.helix", x = df2GramFeatureUmap$word),5] <- "HelixHelix"
df2GramFeatureUmap[grepl(".helix.sheet", x = df2GramFeatureUmap$word),5] <- "HelixSheet"
df2GramFeatureUmap[grepl(".helix.coil", x = df2GramFeatureUmap$word),5] <- "HelixCoil"
```

Also for tsne

```{r}
df2GramFeatureTsne$Config <- "SheetSheet"
df2GramFeatureTsne[grepl(".sheet.coil", x = df2GramFeatureTsne$word),5] <- "SheetCoil"
df2GramFeatureTsne[grepl(".sheet.helix", x = df2GramFeatureTsne$word),5] <- "SheetHelix"
df2GramFeatureTsne[grepl(".coil.coil", x = df2GramFeatureTsne$word),5] <- "CoilCoil"
df2GramFeatureTsne[grepl(".coil.helix", x = df2GramFeatureTsne$word),5] <- "CoilHelix"
df2GramFeatureTsne[grepl(".coil.sheet", x = df2GramFeatureTsne$word),5] <- "CoilSheet"
df2GramFeatureTsne[grepl(".helix.helix", x = df2GramFeatureTsne$word),5] <- "HelixHelix"
df2GramFeatureTsne[grepl(".helix.sheet", x = df2GramFeatureTsne$word),5] <- "HelixSheet"
df2GramFeatureTsne[grepl(".helix.coil", x = df2GramFeatureTsne$word),5] <- "HelixCoil"
```


```{r}
# UMAP
    ggplot(df2GramFeatureUmap, aes(x = x, y = y, color = Config)) +
      geom_point(size = 2) + theme_void() + scale_color_manual(values = c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#a6cee3','#a65628','#f781bf','#999999')) + 
      labs(color = "Secondary structures") + ggtitle("UMAP of the numerical features based on 2 grams") +
      theme(plot.title = element_text(hjust = 0.5)) 
# TSNE
ggplot(df2GramFeatureTsne, aes(x = x, y = y, color = Config)) +
  geom_point(size = 2) + theme_void() + scale_color_manual(values = c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00','#a6cee3','#a65628','#f781bf','#999999'))

# Always looks cool, umap seems to show nice viual info
```

Lets see where the chemical properties cluster

```{r}
# Get the chemical properties for the 2 grams
chemicalPropertiesemb2Grams <- t(as.data.frame(lapply(rownames(embedding2Gram), function(x){
  aminoAcids <- str_split(x, simplify = TRUE, pattern = "")
  aminoAcids <- grep("[A-Z]", aminoAcids, value = T)
  return(c(mean(hydrophobicity[aminoAcids]),  mean(hydrophilicy[aminoAcids]),mean(charge[aminoAcids]), mean(mass[aminoAcids])))
})))

# add names
rownames(chemicalPropertiesemb2Grams) <- rownames(embedding2Gram)
colnames(chemicalPropertiesemb2Grams) <- c("Hydrophobicity", "Hydrophilicy","Charge", "Mass")

# Umap
# Plot
ppp1 <- ggplot(df2GramFeatureUmap, aes(x = x, y = y, label = word, color = chemicalPropertiesemb2Grams[,1])) +
  geom_point(size = 0.9) + theme_void() +
  scale_color_viridis(option = "D", labels=c("Minimum","","", "Maximum")) +
  theme(legend.position = "none") + ggtitle("Hydrophobicity") +
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(color = "")

# Plot
ppp2 <- ggplot(df2GramFeatureUmap, aes(x = x, y = y, label = word,color = chemicalPropertiesemb2Grams[,2])) +
  geom_point(size = 0.9) + theme_void() +
  scale_color_viridis(option = "D")  + theme(legend.position = "none")+ ggtitle("Hydrophilicy") +
  theme(plot.title = element_text(hjust = 0.5))

# Plot
ppp3 <- ggplot(df2GramFeatureUmap, aes(x = x, y = y, label = word,color = chemicalPropertiesemb2Grams[,3])) +
  geom_point(size = 0.9) + theme_void() +
  scale_color_viridis(option = "D")  + theme(legend.position = "none") + ggtitle("Charge") +
  theme(plot.title = element_text(hjust = 0.5))

ppp4 <- ggplot(df2GramFeatureUmap, aes(x = x, y = y, label = word,color = chemicalPropertiesemb2Grams[,4])) +
  geom_point(size = 0.9) + theme_void() +
  scale_color_viridis(option = "D")  + theme(legend.position = "none") + ggtitle("Weight") +
  theme(plot.title = element_text(hjust = 0.5))


ggarrange(ppp1,ppp2, ppp3, ppp4, labels = c("A", "B", "C", "D"), common.legend = TRUE, legend = "right")
```
  
I have no idea how many clusters this might be
We can create a heatmap but with 3600 rows there is not much to see

```{r}
set.seed(2)
Heatmap(embedding2Gram, cluster_rows = F,cluster_columns = F, row_names_gp = gpar(fontsize = 0), show_column_dend = FALSE, show_row_dend = F, km = 1)
```

We can also visualise or even cluster the peptides we have based on the summed
vector representations

```{r}
# Import 1 gram summed features
dataset <- readRDS("/home/jeppe/Dropbox/internship_TNO/Datasets/hemoDatabases/word2vecFeaturesSummed1gram.rds")
# Store the one from the combined (hemo + porter5) dataset
dataset <- dataset[[3]]
# Store hemolytic class and sequences
Y <- dataset$Hemolytic
sequence <- dataset$Sequence
# Get our features per peptide in a matrix
datasetHemoFrom1Grams <- as.matrix(dataset[,1:300])

# Same for 2grams
dataset <- readRDS("/home/jeppe/Dropbox/internship_TNO/Datasets/hemoDatabases/word2vecFeaturesSummed2gram.rds")
dataset <- dataset[[3]]
# Y and sequence are the same as above, no need to store

# Get our features per peptide in a matrix
datasetHemoFrom2Grams <- as.matrix(dataset[,1:300])

rm(dataset)
```

Let's do visualisation of our peptides represented by the summed word2vec representations
from 1grams:

```{r}
set.seed(2)

# Run UMAP
umap1GramPeptides <- umap(datasetHemoFrom1Grams, n_neighbors = 15, n_threads = 2)
# Get in df for ggplot2
df1GramPeptidesUmap  <- data.frame(x = umap1GramPeptides[, 1], y = umap1GramPeptides[, 2], hemo = Y,
                  stringsAsFactors = FALSE)
# Plot UMAP of peptides, colour by hemolytic class
ggplot(df1GramPeptidesUmap, aes(x = x, y = y, color = hemo)) +
  geom_point(size = 2) + theme_void() + scale_color_manual(values = c('#386cb0', '#e41a1c'))

# Run tsne
tsne1GramPeptides <- Rtsne(datasetHemoFrom1Grams, check_duplicates = FALSE)
# Get in df for ggplot2
df1GramPeptidesTsne <- data.frame(x = tsne1GramPeptides[["Y"]][, 1], y = tsne1GramPeptides[["Y"]][, 2], hemo = Y,
                  stringsAsFactors = FALSE)
# Plot TSNE of peptides, colour by hemolytic class
ggplot(df1GramPeptidesTsne, aes(x = x, y = y, color = hemo)) +
  geom_point(size = 2) + theme_void() + scale_color_manual(values = c('#386cb0', '#e41a1c'))
```

We can also try clustering

```{r}
set.seed(2)

# Cluster with km
km1GramPeptides10 <- eclust(datasetHemoFrom1Grams, "kmeans", iter.max = 25, k = 10)
km1GramPeptides15 <- eclust(datasetHemoFrom1Grams, "kmeans", iter.max = 25, k = 15)
km1GramPeptides20 <- eclust(datasetHemoFrom1Grams, "kmeans", iter.max = 25, k = 20)

# Show clusters in umap
ggplot(df1GramPeptidesTsne, aes(x = x, y = y, color = as.factor(km1GramPeptides10[["cluster"]]))) +
  geom_point() 

ggplot(df1GramPeptidesTsne, aes(x = x, y = y, color = as.factor(km1GramPeptides15[["cluster"]]))) +
  geom_point() 

ggplot(df1GramPeptidesTsne, aes(x = x, y = y, color = as.factor(km1GramPeptides20[["cluster"]]))) +
  geom_point() 
```

```{r}
# Clusters with dbscan
kNNdistplot(as.matrix(datasetHemoFrom1Grams), k = 305)
dbscan1gram <- dbscan(as.matrix(datasetHemoFrom1Grams), eps = 250, minPts = 305)

# Umap
ggplot(df1GramPeptidesTsne, aes(x = x, y = y, color = as.factor(dbscan1gram$cluster))) +
  geom_point() 
```

Lets do the same for the peptides represented by the summed vectors of the 
2-grams.

```{r}
set.seed(2)

# Run UMAP
umap2GramPeptides <- umap(datasetHemoFrom2Grams, n_threads = 2)
# Get in df for ggplot2
df2GramPeptidesUmap  <- data.frame(x = umap2GramPeptides[, 1], y = umap2GramPeptides[, 2], hemo = Y,
                  stringsAsFactors = FALSE)
# Plot UMAP of peptides, colour by hemolytic class
ggplot(df2GramPeptidesUmap, aes(x = x, y = y, color = hemo)) +
  geom_point(size = 2) + theme_void() + scale_color_manual(values = c('#386cb0', '#e41a1c'))

# Run tsne
tsne2GramPeptides <- Rtsne(datasetHemoFrom2Grams, check_duplicates = FALSE)
# Get in df for ggplot2
df2GramPeptidesTsne <- data.frame(x = tsne2GramPeptides[["Y"]][, 1], 
                                  y = tsne2GramPeptides[["Y"]][, 2], hemo = Y,
                  stringsAsFactors = FALSE)
# Plot TSNE of peptides, colour by hemolytic class
ggplot(df2GramPeptidesTsne, aes(x = x, y = y, color = hemo)) +
  geom_point(size = 2) + theme_void() + scale_color_manual(values = c('#386cb0', '#e41a1c'))
```

Clustering

```{r}
set.seed(2)

# Cluster with km
km2GramPeptides10 <- eclust(datasetHemoFrom2Grams, "kmeans", iter.max = 25, k = 10)
km2GramPeptides15 <- eclust(datasetHemoFrom2Grams, "kmeans", iter.max = 25, k = 15)
km2GramPeptides20 <- eclust(datasetHemoFrom2Grams, "kmeans", iter.max = 25, k = 20)

# Show clusters in umap
ggplot(df2GramPeptidesUmap, aes(x = x, y = y, color = as.factor(km2GramPeptides10[["cluster"]]))) +
  geom_point() 

ggplot(df2GramPeptidesUmap, aes(x = x, y = y, color = as.factor(km2GramPeptides15[["cluster"]]))) +
  geom_point() 

ggplot(df2GramPeptidesUmap, aes(x = x, y = y, color = as.factor(km2GramPeptides20[["cluster"]]))) +
  geom_point() 
```

This does not really seem to add anything of value
