```{r}
library(ranger)
library(caret)
library(dplyr)
library(stringr)
library(parallel)
library(foreach)
library(protr)
library(doParallel)
```

Here we will test our RF + peptide composition model on some extra peptides
I extracted from the internet and cross reverenced with the hemo dataset to see
if they werent in there. I also took intrachainbond and N/C-term mod info from 
the papers these peptides were in and put them in the hemo or non hemo class
based on the system used by HAPPENN.

Import the peptides
```{r}
extraTest <-read.csv("file:///home/jeppe/Dropbox/internship_TNO/Datasets/Extra_test/extraTest.csv")
```

I already predicted the Secondary structure, we can extract it and bind it

```{r}
# List the files with secondary strcuture information
SS3_files <- list.files(path = "/home/jeppe/Dropbox/internship_TNO/Datasets/SS3/extraTest_ss3/", pattern = ".+ss3", full.names = TRUE)

# Lets build a function to extract secondary strcuture info 
# input is vector of .ss3 file (ourput from Porter5) locations
extractSS3 <- function(files){ 

# Need this to dissect .ss3 file; C coil, H helix, E sheet
ss3Structures <- c("C", "H", "E") 

# For each file in the files list
ss3 <- foreach(file = files, .combine = rbind) %do% {
# Read porter 5 output as table  
ss3Output <- read.table(file = file, header = F)
# Take sequences, which are column 2, transpose and make vector and collapse
ss3Sequence <- paste(as.vector(t(ss3Output[2])), collapse = "") 
# Take secondary structure info, which are column 3, transpose and make vector 
# and collapse
ss3Results <- as.vector(t(ss3Output[3])) #takes results
# The summary function is used to quickly get the amount of occurances of the 
# three classes for each sequence, then this is devided by the sequence length 
# to get the percentage of occurance. It is then stored as a df with three columns
ss3 <- t(as.data.frame(x = summary(factor(ss3Results, levels = ss3Structures), maxsum = 4)/length(ss3Results), row.names = c("Coil", "Helix", "Sheet"))) 
ss3 <- as.data.frame(ss3)
# Convenient to have the sequence
rownames(ss3) <- ss3Sequence 
# And the secondary structure in sequence format
ss3$SS <- paste(ss3Results, collapse = "")
return(ss3)}

as.data.frame(ss3)
}
# Run our function 
SS3 <- extractSS3(SS3_files)
```

Bind ssc info

```{r}
SS3$Sequence <- rownames(SS3) 

l_temp <- length(extraTest$Sequence) #to check if you have all SS3 and see if you lose rows with merge
extraTest <-
    merge(extraTest, SS3[,c(1:3,5)], by = "Sequence") #merge, don't include the full secondary structure sequence
l_temp_2 <-
    length(extraTest$Sequence)
cat("This should be true, if it is not you miss ss3 data:  ")
!(l_temp_2 < l_temp)
```

Add AAC and DC info
```{r}
# Frequency of 20 single aminoacids
AAC <- as.data.frame(t(sapply(as.vector(extraTest$Sequence), extractAAC))) 
# Frequency of 20*20 = 400 aminoacid combinations
DC <- as.data.frame(t(sapply(as.vector(extraTest$Sequence), extractDC)))

# Combine
extraTest <- cbind(extraTest, AAC, DC)
```

Load best model, modify df a bit

```{r}
# Load model we determined was the best
bestRFModel <- readRDS("/home/jeppe/Dropbox/internship_TNO/R/2_RF/RF_models/finalModel/bestRFModel.rds")

# Modify df a bit
colnames(extraTest)[colnames(extraTest) == 'NA'] <- "NA." 
# STore and remove sequences
extraTestSequences <-extraTest$Sequence
extraTest$Sequence <- NULL
# And our y column has to contain factors for the random forest library to work
extraTest$Hemolytic <- as.factor(make.names(factor(extraTest$Hemolytic)))
extraTest <- extraTest %>% mutate_if(is.character, as.factor) 
# Set correct order for positive and negative in performance measuers
extraTest$Hemolytic <-factor(extraTest$Hemolytic, levels = rev(levels(extraTest$Hemolytic)))
```

Predict and get performance measures, bind sequences

```{r}
pred <- predict(bestRFModel, extraTest)
confusionMatrix(as.factor(pred$predictions),
                    as.factor(extraTest$Hemolytic),
                    positive = "X1")

pred1 <- data.frame("Predictions" = pred$predictions, "Sequence" = extraTestSequences)
```

Compare to predictions of HAPPENN 
https://research.timmons.eu/
and HemoPred http://codes.bio/hemopred/

```{r}
# Import their predictions
predHAPPENN <- read.csv("file:///home/jeppe/Dropbox/internship_TNO/R/7_extraTestPeptides/happennExtraTestPredictions.csv", header = F)
predHemoPred <- read.csv("file:///home/jeppe/Dropbox/internship_TNO/R/7_extraTestPeptides/HLPpred-FuseExtraTestPredictions.csv", header = F)[1:2]

# Add colnames
colnames(predHAPPENN) <- c("Sequence", "predHAPPENN")
colnames(predHemoPred) <- c("Sequence", "predHLPpred-Fuse")

# Join
pred2 <- inner_join(predHAPPENN, predHemoPred, by = "Sequence")
```

Make df so we can easily compare

```{r}
# Join predictions
df <- inner_join(pred1, pred2)
# Add truth and names 
truthNames <-read.csv("file:///home/jeppe/Dropbox/internship_TNO/Datasets/Extra_test/extraTest.csv")
truthNames <- truthNames[c(1:2,6)]
# make class names same
truthNames$Hemolytic <- ifelse(truthNames$Hemolytic == 1, "X1", "X0")

df <- inner_join(truthNames, df)

write.csv(df, "/home/jeppe/Dropbox/internship_TNO/R/7_extraTestPeptides/predictionResultsExtraTest.csv")
```

One last thing, what if we use all peptides to train a RF and classify the extra
test case?

```{r}
hemo_df <- read.csv("/home/jeppe/Dropbox/internship_TNO/Datasets/Hemo_databases/hemo_df_ICB_NCtermMods_SS3_AAC_DC_18july.csv")
# Remove some columns that we dont need
hemo_df$X <- NULL
hemo_df$Source <- NULL
hemo_df$Sequence <- NULL

# Edit the colnames of the NA dipeptide, Ranger does not like this.
colnames(hemo_df)[colnames(hemo_df) == 'NA'] <- "NA." 
# And our y column has to contain factors for the random forest library to work
hemo_df$Hemolytic <- as.factor(make.names(factor(hemo_df$Hemolytic)))
hemo_df <- hemo_df %>% mutate_if(is.character, as.factor) 
# Set correct order for positive and negative in performance measuers
hemo_df$Hemolytic <-factor(hemo_df$Hemolytic, levels = rev(levels(hemo_df$Hemolytic))) 

# Weights
train_weights <- ifelse(hemo_df$Hemolytic == "X1",
                        (1/table(hemo_df$Hemolytic)[1]) * 0.5,
                        (1/table(hemo_df$Hemolytic)[2]) * 0.5)
# Train rf
bestRFModel <- ranger(Hemolytic  ~ ., data = hemo_df, num.trees = 2000, min.node.size = 1, mtry = 23, case.weights = train_weights)

predFullDataset <- predict(bestRFModel, extraTest)
confusionMatrix(as.factor(predFullDataset$predictions),
                    as.factor(extraTest$Hemolytic),
                    positive = "X1")
# STore in df
predFullDataset <- data.frame("PredictionsFullDS" = pred$predictions, "Sequence" = extraTestSequences)

# bind to compare between rf on train and rf on full
df <- inner_join(df, predFullDataset)

# predictions are the same
```

